---
title: "NHANES explorer"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: https://github.com/rstudio-education/shiny-rmed18/blob/master/01-flexdash/nhanesdash.Rmd
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(tidyverse)
library(NHANES)
library(DT)
```

Sidebar {.sidebar}
=======================================================================

### NHANES

Filter the data for characteristics you're interested in studying:

```{r}
hr()

checkboxGroupInput(inputId = "survey_year", label = "Survey year",
                   choices = c("2011 - 2012" = "2011_12", 
                               "2009 - 2010" = "2009_10"),
                   selected = "2011_12")

hr()

sliderInput(inputId = "age", label = "Age range",
            min = min(NHANES$Age), max = max(NHANES$Age),
            value = c(18, 55))

hr()

checkboxGroupInput(inputId = "education", label = "Education level",
                   choices = levels(NHANES$Education),
                   selected = c("Some College", "College Grad"))

sel_data <- reactive({
  NHANES %>%
    filter(
      SurveyYr %in% input$survey_year,
      between(Age, input$age[1], input$age[2]),
      Education %in% input$education
      ) 
})
```

Dashboard
=======================================================================

Row
-----------------------------------------------------------------------

### Percentage of NHANES sample {.value-box}

```{r}
renderGauge({
  p <- round( nrow(sel_data()) / nrow(NHANES) * 100, 0)
  gauge(p, min = 0, max = 100, symbol = "%", 
        gaugeSectors(success = c(80, 100), warning = c(40, 79), 
                     danger = c(0, 39)))
})
```

### Average age {.value-box}

```{r}
renderValueBox({
  valueBox(value = round(mean(sel_data()$Age), 2), icon = "glyphicon-time")
})
```

### Percentage of smokers {.value-box}

```{r}
perc_smoker <- reactive({
  sel_data() %>%
    filter(!is.na(SmokeNow)) %>%
    count(SmokeNow) %>%
    mutate(p = n / sum(n)) %>% 
    slice(2) %>% 
    pull(p)
})
renderValueBox({
  valueBox(value = scales::percent(perc_smoker()), 
           icon = "glyphicon-thumbs-down")
})
```

Row
-----------------------------------------------------------------------

### Alcohol consumption - plot {data-width=700}

The x-axis in the plot below shows the estimated number of days over the past 
year that the participants drank alcoholic beverages. This variable is reported for 
participants aged 18 years or older.

Examine the relationship between alcohol consumption and ...

```{r}
selectInput(inputId = "x", label = "",
            choices = c("Trouble sleeping" = "SleepTrouble",
                        "Marijuana use" = "Marijuana", 
                        "Hard drug use" = "HardDrugs",
                        "Physically active" = "PhysActive"),
            selected = "SleepTrouble")

xlab <- reactive({
  case_when(
    input$x == "SleepTrouble" ~ "Trouble sleeping",
    input$x == "Marijuana" ~ "Marijuana use",
    input$x == "HardDrugs" ~ "Hard drug use",
    input$x == "PhysActive" ~ "Physically active"
  )
})
```


```{r}
renderPlot({
  ggplot(data = sel_data(), 
         mapping = aes_string(x = input$x, y = "AlcoholYear")) +
    geom_boxplot() +
    coord_flip() +
    labs(x = xlab(), y = "")}, 
  height = 275, width = 650)
```

### Alcohol consumption - summary {data-width=340}

```{r}
group_vars <- reactive({ syms(input$x) })

renderTable({
  sel_data() %>%
    group_by( !!!group_vars() ) %>%
    summarise(Mean = mean(AlcoholYear, na.rm = TRUE),
              Median = median(AlcoholYear, na.rm = TRUE))
})
```

Data
=======================================================================

### About the data

This [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) uses data 
from the American National Health and Nutrition Examination surveys (NHANES). 
Specifically, the data come from the [NHANES package ](https://cran.r-project.org/web/packages/NHANES/), which provides 
modified versions of NHANES data for educational purposes. The package author 
cautions that the package is "NOT suitable for use as a research database".

The data used in the analysis is shown below.

```{r}
renderDataTable({
  datatable(
    sel_data() %>%
      select(AlcoholYear, SleepTrouble, Marijuana, HardDrugs, PhysActive)
    )
})
```

